---
title: "Two Way Anova"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(GGally)    # for ggpairs()
library(tidyverse) # for read_csv
library(FSA)       # for Summarize()
library(car)       # for Anova()
library(rcompanion)# for plotNormalHistogram()
library(multcompView)# for ...
library(lsmeans)   #
library(multcomp)  # for cld
```

## Two Way ANOVA
Here is where we will do Two Way ANOVA, with League and Division as the two categorical variables.

```{r}
# will replace this when master .rmd file does transform.
hitters <- read_csv("Hitters.csv")
hitters$Division <- as.factor(hitters$Division)
hitters$League   <- as.factor(hitters$League)
for_anova <- hitters %>% filter(!is.na(Salary)) %>% select(Salary,Division,League) %>%
  mutate(logs = log(Salary))
```

### Hypotheses

### Check Assumptions

#### Are the cell sizes balanced

```{r}
Summarize(logs ~ League + Division,data=for_anova)
```

```{r}
for_anova %>% ggpairs()
```

```{r}
ggplot(hitters,aes(x=League,y=Salary,group=Division,colour=Division)) + ylab("Salary") + xlab("League") + geom_point()
```

```{r}
hitters_aov.fit<-aov(logs~Division+League+Division:League,data=for_anova)
par(mfrow=c(1,2))
plot(hitters_aov.fit$fitted.values,hitters_aov.fit$residuals,ylab="Resdiduals",xlab="Fitted")
qqnorm(hitters_aov.fit$residuals)
```

Define a linear model

```{r}
two_way_anova_model = lm(logs ~ Division + League + Division:League,data=for_anova)
summary(two_way_anova_model)
```

Conduct Analysis of Variance

```{r}
Anova(two_way_anova_model,type="III")
```

Histogram of Residuals

```{r}
two_way_anova_residuals <- residuals(two_way_anova_model)
plotNormalHistogram(two_way_anova_residuals)
```

Scatterplot of Residuals
```{r}
plot(fitted(two_way_anova_model),two_way_anova_residuals)
```

LS means for main effects (Division)
```{r}
marginal <- lsmeans(two_way_anova_model,~Division)
pairs(marginal,adjust="tukey")
CLD <- cld(marginal)
CLD
```

LS means for main effects (League)
```{r}
marginal <- lsmeans(two_way_anova_model,~League)
pairs(marginal,adjust="tukey")
CLD <- cld(marginal)
CLD
ggplot(CLD,aes(x=League,y=lsmean,label=.group)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower.CL,ymax=upper.CL)) +
  ylab("Least square mean\nLog Salary")
```

LS means for main effects (Division)
```{r}
marginal <- lsmeans(two_way_anova_model,~Division)
pairs(marginal,adjust="tukey")
CLD <- cld(marginal)
CLD
ggplot(CLD,aes(x=Division,y=lsmean,label=.group)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower.CL,ymax=upper.CL)) +
  ylab("Least square mean\nLog Salary")
```


LS means for Interactiopn
```{r}
marginal <- lsmeans(two_way_anova_model,~Division+League)
pairs(marginal,adjust="tukey")
CLD <- cld(marginal)
CLD
pd = position_dodge(0.6)
ggplot(CLD,aes(x=Division,y=lsmean,colour=League,label=.group)) +
  geom_point(position=pd) +
  geom_errorbar(aes(ymin=lower.CL,ymax=upper.CL,position=pd)) +
  ylab("Least square mean\nLog Salary")
```


