---
title: "Two Way Anova"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(GGally)    # for ggpairs()
library(tidyverse) # for read_csv
library(FSA)       # for Summarize()
library(car)       # for Anova()
library(rcompanion)# for plotNormalHistogram()
library(multcompView)# for ...
library(lsmeans)   #
library(multcomp)  # for cld
```

## Two Way ANOVA
Here is where we will do Two Way ANOVA, with League and Division as the two categorical variables.

```{r}
# will replace this when master .rmd file does transform.
hitters <- read_csv("Hitters.csv")
hitters$Division <- as.factor(hitters$Division)
hitters$League   <- as.factor(hitters$League)

for_anova <- hitters %>% dplyr::filter(!is.na(Salary)) %>% dplyr::select(Salary,Division,League) %>%
  dplyr::mutate(logs = log(Salary))
```

### Hypotheses

For The Division:
Hl0: The mean of the log salary is the same between East and West.
Hla: The mean of the log salary is different between East and West.

For The League:
Hd0: The mean of the log salary is the same between American and National
Hda: The mean of the log salary is different between American and National

For Interaction:


### Check Assumptions

The populations from which the samples were obtained must be normally or approximately normally distributed.
The samples must be independent.
The variances of the populations must be equal.
The groups must have the same sample size.

This summary shows that the groups have very nearly the same sample size.  The standard deviations are also nearly the same.  

```{r}
Summarize(logs ~ League + Division,data=for_anova)
```

These plots of histograms and box plots help confirm that the variances are the same.

```{r}
for_anova %>% dplyr::select(logs,Division,League) %>% ggpairs()
```


```{r}
hitters_aov.fit<-aov(logs~Division+League+Division:League,data=for_anova)
summary(hitters_aov.fit)
```

```{r}
par(mfrow=c(1,2))
plot(hitters_aov.fit$fitted.values,hitters_aov.fit$residuals,ylab="Resdiduals",xlab="Fitted")
qqnorm(hitters_aov.fit$residuals)
```

Define a linear model

```{r}
two_way_anova_model = lm(logs ~ Division + League + Division:League,data=for_anova)
summary(two_way_anova_model)
```

Conduct Analysis of Variance

```{r}
Anova(two_way_anova_model,type="III")
```

Histogram of Residuals

```{r}
two_way_anova_residuals <- residuals(two_way_anova_model)
plotNormalHistogram(two_way_anova_residuals)
```

Scatterplot of Residuals
Note: jitter is added to help visualize the points.  There are only four truly distinct 
fitted values.
```{r}
plot(fitted(two_way_anova_model),two_way_anova_residuals)
resi_df = data.frame(cbind(fitted=fitted(two_way_anova_model),resid=two_way_anova_residuals))
resi_df %>% ggplot(aes(x=fitted,y=resid)) + geom_point() + geom_jitter(width=0.015,height=0.0)
```

LS means for main effects (Division)
```{r}
marginal <- lsmeans(two_way_anova_model,~Division)
pairs(marginal,adjust="tukey")
CLD <- cld(marginal)
CLD
ggplot(CLD,aes(x=Division,y=lsmean,label=.group)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower.CL,ymax=upper.CL)) +
  ylab("Least square mean\nLog Salary")
```

LS means for main effects (League)
```{r}
marginal <- lsmeans(two_way_anova_model,~League)
pairs(marginal,adjust="tukey")
CLD <- cld(marginal)
CLD
ggplot(CLD,aes(x=League,y=lsmean,label=.group)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower.CL,ymax=upper.CL)) +
  ylab("Least square mean\nLog Salary")
```


LS means for Interactiopn
```{r}
marginal <- lsmeans(two_way_anova_model,~Division+League)
pairs(marginal,adjust="tukey")
CLD <- cld(marginal)
CLD
pd = position_dodge(0.6)
ggplot(CLD,aes(x=Division,y=lsmean,colour=League,label=.group)) +
  geom_point(position=pd) +
  geom_errorbar(aes(ymin=lower.CL,ymax=upper.CL),position=pd) +
  ylab("Least square mean\nLog Salary")
```


